---
title: "p8105_hw3_yl2625"
author: "Yurou Liu"
date: "2025-10-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
load in data

```{r}
library(dplyr)
library(p8105.datasets)

data("instacart") 
insta_ez_to_read = instacart %>% 
  select(aisle, everything())
```

### 1-1 How many aisles are there, and which aisles are the most items ordered from?

```{r}
# group by same aisles and sum up the order number for the same group 
total_aisle_df = insta_ez_to_read %>% 
  group_by(aisle) %>% 
  summarise(total_orders = n(), aisle = first(aisle)) %>%
  arrange(desc(total_orders))

# count the number of different aisles
num_aisles = n_distinct(total_aisle_df$aisle)

# find out the best seller
best_seller = total_aisle_df %>% 
  # pick out the aisle with the most orders, of there are aisles with the same number of orders, they will be picked out together
  filter(total_orders == max(total_orders))

print(num_aisles)
print(best_seller)
```
According to the printed results, there are 134 different aisles, and the most items are ordered from fresh vegetables, with 150609 orders in total.

### 1-2 make a plot

```{r, fig.height=16, fig.width=10}
plot_df = total_aisle_df %>% 
  filter(total_orders > 10000)

library(ggplot2)

plot_df %>% 
  ggplot(aes(y = reorder(aisle, total_orders), x = total_orders)) + 
  geom_col() +
  labs(title = "Aisles with 10000+ Orders", x = "Number of Orders", y = "Aisle") + 
  # add order number labels to the columns.
  geom_text(aes(label = format(total_orders, big.mark = ",")), hjust = -.05, size = 3) + 
  # expand x axis from the right side.
  scale_x_continuous(expand = expansion(mult = c(0, .1))) + 
  theme(plot.title = element_text(hjust = .5))
```

### 1-3 Table: three most popular items in 'baking ingredients', 'dog food care' and 'packaged vegetables fruits' with number of orders.

```{r}
df_for_table = insta_ez_to_read %>% 
  select(aisle, product_name) %>% 
  group_by(product_name) %>% 
  summarise(total_orders = n(), 
            aisle = first(aisle)) %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits" )) %>% 
  group_by(aisle) %>% 
  slice_max(total_orders, n = 1) %>% 
  ungroup() %>% 
  select(aisle, everything()) %>% 
  print()
```

### 1-4 Table: mean hour of the day at which pink lady apples and coffee ice cream are ordered on each day of the week.

```{r}
library(tidyr)

df_for_table2 = insta_ez_to_read %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name) %>% 
  mutate(row_id = row_number()) %>% 
  pivot_wider(names_from = product_name, 
              values_from = order_hour_of_day) %>% 
  select(-row_id) %>% 
  arrange(order_dow) %>% 
  group_by(order_dow) %>% 
  summarise(mean_order_hour_apples = mean(`Pink Lady Apples`, na.rm = TRUE), 
         mean_order_hour_icecream = mean(`Coffee Ice Cream`, na.rm = TRUE)) %>% 
  mutate(order_dow = recode(order_dow, 
                            `0` = "Sunday", 
                            `1` = "Monday", 
                            `2` = "Tuesday", 
                            `3` = "Wednesday", 
                            `4` = "Thursday", 
                            `5` = "Friday", 
                            `6` = "Saturday")) %>% 
  print()
```


## Problem 2

```{r}
zillow_df = readr::read_csv("./data/zip_nyc.csv") %>% 
  janitor::clean_names()

a_zillow_df = zillow_df %>% 
  pivot_longer(c("x2015_01_31":"x2024_08_31"), 
               names_to = "date", 
               names_prefix = "x")

zip_count = a_zillow_df %>% 
  filter(!is.na(value)) %>%
  group_by(region_name) %>% 
  summarise(count = n())

```


### 2-1-a How many ZIP codes are observed 116 times?

```{r}
obs_116_times = zip_count %>% 
  filter(count == 116) %>% 
  print()
```

According to the printed results, there are 48 zip codes that were observed 116 times in total.

### 2-1-b How many are observed fewer than 10 times?

```{r}
obs_less_than_10 = zip_count %>% 
  filter(count < 10) %>% 
  print()
```

According to the printed results, there are 26 zip codes that were observed less than 10 times.

### 2-1-c Why are some ZIP codes are observed rarely and others observed in each month?

According to a_zillow_df, differences in zip code observation times may due to the various data collection starting time. For example, 11368 was first observed from 2023-01-31, which may imply that the data collection in this area began later. Besides, there is also possibility that some zip codes were merged or split due to administrative changes, which can lead to the discontinuity and incomplete of these historical data.

### 2-2-a Table: average rental price in wach borough and year. 

```{r}
queens_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Queens County") %>% 
  group_by(year) %>% 
  summarise(ave_rental_prices = mean(value, na.rm = TRUE))

kings_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Kings County") %>% 
  group_by(year) %>% 
  summarise(ave_rental_prices = mean(value, na.rm = TRUE))

richmond_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Richmond County") %>% 
  group_by(year) %>% 
  summarise(ave_rental_prices = mean(value, na.rm = TRUE))

bronx_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Bronx County") %>% 
  group_by(year) %>% 
  summarise(ave_rental_prices = mean(value, na.rm = TRUE))

ny_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "New York County") %>% 
  group_by(year) %>% 
  summarise(ave_rental_prices = mean(value, na.rm = TRUE))

p22a_df = bind_rows(
  queens_df %>% mutate(county = "Queens"), 
  kings_df %>% mutate(county = "Kings"), 
  richmond_df %>% mutate(county = "Richmond"),
  bronx_df %>% mutate(county = "Bronx"),
  ny_df %>% mutate(county = "New York")) %>% 
  pivot_wider(names_from = county, 
              values_from = ave_rental_prices) %>% 
  print()
```

### 2-2-b Comment on the table

Overall speaking, all boroughs show an increase in rental prices from 2015 to 2024. The New York County stays the most expensive throughout the whole 10 years, followed by Kings, Queens, Richmond and Bronx. Rental prices in Queens and Kings County decreased from 2019 to 2021, rental prices in New York County decreased sharply from 2019 to 2020, probably due to the COVID-19 pandemic. While rental price in Richmond County increased from 2020 to 2024 and rental price of Bronx County kept increasing throughout the whole period observed.

### 2-3-a

```{r, fig.height=16, fig.width=20}
queens_23a_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Queens County") %>% 
  group_by(region_name, year) %>% 
  summarise(county = "Queens", ave_rental_prices = mean(value, na.rm = TRUE))

kings_23a_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Kings County") %>% 
  group_by(region_name, year) %>% 
  summarise(county = "Kings", ave_rental_prices = mean(value, na.rm = TRUE))

richmond_23a_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Richmond County") %>% 
  group_by(region_name, year) %>% 
  summarise(county = "Richmond", ave_rental_prices = mean(value, na.rm = TRUE))

bronx_23a_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Bronx County") %>% 
  group_by(region_name, year) %>% 
  summarise(county = "Bronx", ave_rental_prices = mean(value, na.rm = TRUE))

ny_23a_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "New York County") %>% 
  group_by(region_name, year) %>% 
  summarise(county = "New York", ave_rental_prices = mean(value, na.rm = TRUE))

p23a_df = bind_rows(kings_23a_df, queens_23a_df, richmond_23a_df, bronx_23a_df, ny_23a_df)

p23a_plot = p23a_df %>% 
  filter(!is.na(ave_rental_prices)) %>% 
  ggplot(aes(x = year, 
             y = ave_rental_prices, 
             group = region_name)) + 
  geom_line(color = "#52C0EF") + 
  geom_point(color = "#52C0EF") + 
  geom_text(data = . %>% 
              group_by(region_name) %>% 
              filter(!is.na(ave_rental_prices)) %>% 
              slice(1), 
            aes(label = region_name), 
            hjust = 1.1,
            size = 2.5, 
            color = "black",
            check_overlap = TRUE) + 
  facet_wrap(. ~ county, nrow = 2, scales = "free_y") +
  labs(title = "Average NYC Rental Prices Within ZIP Codes for All Available Years",
       y = "Average Rental Prices", 
       x = "Year") + 
  theme(legen.position = "none", 
        plot.title = element_text(hjust = .5, size = 20))
```

### 2-3-b Comment on the plot

The plot shows the overall changes in the rental prices of regions of different zip codes in the five counties. All boroughs exhibit upward trends, particularly during the 2021-2024 post-pandemic recovery period. There also exists intra-borough disparities. Rental prices in Bronx County demonstrate steep percentage growth rates in recent years. 

### 2-4-a

```{r, fig.height=16, fig.width=20}
queens_24a_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Queens County",
         year == "2023")

kings_24a_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Kings County",
         year == "2023")

richmond_24a_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Richmond County",
         year == "2023")

bronx_24a_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "Bronx County",
         year == "2023")

ny_24a_df = a_zillow_df %>% 
  separate(date, c("year", "month","day")) %>% 
  filter(county_name == "New York County",
         year == "2023")

p24a_df = bind_rows(queens_24a_df, kings_24a_df, richmond_24a_df, bronx_24a_df, ny_24a_df)

p24a_plot = p24a_df %>% 
  ggplot(aes(x = month, 
           y = value, 
           group = region_name)) +
  geom_line(color = "#5EE14D") + 
  geom_point(color = "#5EE14D") + 
  geom_text(data = . %>% 
              group_by(region_name) %>% 
              filter(!is.na(value)) %>% 
              slice(1),
            aes(label = region_name), 
            hjust = 1.1, 
            size = 2.5, 
            color = "black", 
            check_overlap = TRUE) + 
  facet_wrap(. ~ county_name, nrow = 2, scales = "free_y") + 
  labs(title = "Average Rental Price Within Each ZIP Code Over Each Month in 2023",
       x = "Month", 
       y = "Rental Prices") + 
  theme(plot.title = element_text(hjust = .5, size = 20))
```

### 2-4-b Comment on the plot

This plot shows variations across boroughs in NYC in 2023. Counties such as Queens and Kings County show higher prices and larger fluctuations compared to the more stable rental prices in Bronx and Richmond Counties. Within the same county, there are also intra-county disparities in rental prices. Several zip codes exhibit notable volatility, which suggest dynamics of the market and others maintain relatively steady trajectories throughout the year.

### 2-5 Combine the previous two plots and export to 'results' folder.

```{r}
library(patchwork)

combined_plot = p23a_plot / p24a_plot
ggsave("./results/combined_plot.png", 
       plot = combined_plot, 
       width = 20, 
       height = 32, 
       dpi = 300)
```

## Problem 3

### 3-1 Data tidying

```{r}
accel_df = readr::read_csv("./data/nhanes_accel.csv") %>% 
  janitor::clean_names()
covar_df = readr::read_csv("./data/nhanes_covar.csv", skip = 4) %>% 
  janitor::clean_names() 

merged_df <- left_join(accel_df, covar_df, by = "seqn") %>% 
  select(sex, age, bmi, education, everything()) %>% 
  filter(age >= 21, 
         !is.na(sex),
         !is.na(bmi), 
         !is.na(education)) %>% 
  mutate(sex = factor(sex, 
                      levels = c(1, 2), 
                      labels = c("male", "female")), 
         education = factor(education, 
                            levels = c(1, 2, 3), 
                            labels = c("Less than high school", 
                                       "High school equivalent", 
                                       "More than high school"), 
                            ordered = TRUE))
```

### 3-2-a Table: number of men and women in each education category

```{r}
p32a_df = merged_df %>% 
  select(sex, education)  %>% 
  group_by(education, sex) %>%
  summarise(n = n(), .groups = "drop") %>% 
  pivot_wider(names_from = education, 
              values_from = n) %>% 
  print()
```

### 3-2-b Plot: age distributions for men and women in each education category.

```{r}
p32b_df = merged_df %>% 
  select(education, sex, age)

p32b_plot = p32b_df %>% 
  ggplot(aes(x = age, 
             fill = sex)) + 
  geom_histogram(color = "black") +
  facet_wrap(. ~ education) + 
  scale_x_continuous(breaks = seq(20, 80, by = 5)) + 
  labs(title = "Age Distributions for Men and Women in Each Education Category", 
       x = "Age", 
       y = "Count") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6.5))
```

### 3-2-c Comment on the table and plot

Table: As shown in p32a_df, male has a highest percentage in the education level "high school equivalent", and female has a highest percentage in the education level "more than high school". The distribution of each gender in each education level is relatively balanced. 

Plot: a large percentage of the data is from users who has an education level that is more than high school. In the education level which is high school equivalent, age interval 30-40, 50-60, and 65-70, number of male is far more than female. In the education level which is less than high school, there exits blank area in histogram, this may be caused by the exclusion of rows with na values in demographic data, or also can be caused by sampling bias.

### 3-3-a Plot: total activity by age, sex and education level.

```{r}
p33a_df = merged_df %>% 
  mutate(activity = rowSums(across(min1:min1440)), na.rm = TRUE) %>% 
  select(sex, age, education, activity) 

p33a_plot = p33a_df %>%
  ggplot(aes(x = age, 
             y = activity)) +
  geom_point(size = .8, 
             alpha = .6, 
             aes(color = sex, fill = sex) ) + 
  geom_smooth(aes(color = sex), 
              se = FALSE) + 
  labs(title = "Total Activity by Age, Sex and Education Level",
       x = "Age", 
       y = "Total Activity") + 
  scale_x_continuous(breaks = seq(20, 80, by = 5)) + 
  facet_wrap(. ~ education) + 
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 6.5),
        plot.title = element_text(hjust = .5))
```

### 3-3-b Comment on the plot

According to the p33a_plot, of all three education levels, the total activity has an decreasing overall trend. There exist fluctuations in total activity from age 20 to age 80 participant, which may due to different perspectives in the living style among different ages and education levels. Under the education level which is mare than high school, total activity of male of every observed age is higher than female.

### 3-4-a Plot: Mean hourly activity by hour, sex and education level.

```{r}
p34a_df = merged_df %>% 
  pivot_longer(cols = starts_with("min"), 
               names_to = "minute", 
               values_to = "activity") %>% 
  mutate(minute_num = as.numeric(gsub("min", "", minute)), 
         hour = floor((minute_num - 1) / 60)) %>% 
  group_by(seqn, education, sex, hour) %>%
  summarise(hourly_activity_per_person = sum(activity, na.rm = TRUE), 
            .groups = "drop") %>%
  group_by(education, sex, hour) %>% 
  summarise(mean_hourly_activity = mean(hourly_activity_per_person, 
            na.rm = TRUE), 
            .groups = "drop")

p34a_plot = p34a_df %>% 
  ggplot(aes(x = hour, 
             y = mean_hourly_activity,)) + 
  geom_point(aes(fill = sex, 
                 color = sex), 
             size = .8, 
             alpha = .6) + 
  geom_smooth(se = FALSE, 
              aes(color = sex)) + 
  facet_wrap(. ~ education) + 
  labs(x = "Hour", 
       y = "Mean Hourly Activity", 
       title = "Mean Hourly Activity by Hour, Sex and Education Level") + 
  scale_x_continuous(breaks = seq(0, 24, by = 4)) +
  theme(plot.title = element_text(hjust = .5), 
        axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 8))
  
```

### 3-4-b Comment on the plot

According to p34a_plot, in all three education levels and both two genders, the overall trend of mean hourly activity level in a day first increases and arrives the peak at around 12pm and then decreases. Difference in mean hourly activity level among different education levels is not too much. And it can also be noticed that in the beginning of the day, mean hourly activity level of male is higher than female. Under "equivalent to high school" and "more than high school" education level, the mean hourly activity level of female is higher than male almost in every hour. 